---
title: "BSR Exp 2 Analysis"
author: "S.C. Cook"
date: "October 25, 2018"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# Package import


```{r, message = FALSE, warning = FALSE}
require(tidyverse)
```


# Data import


```{r}
master <- read.csv(file = "../Data/BSR_exp_2.csv", header = T, sep = ",")
```


# Get in tidy format

```{r}
n2_long <- master %>%
  filter(ANALYTE == "N2")

o2_long <- master %>%
  filter(ANALYTE == "O2")
```


# Exploratory plotting

```{r}
n2_histo <- ggplot() +
  geom_histogram(data = n2_long,
                 aes(x = CONC, fill = TREATMENT),
                 binwidth = 1) +
  theme_classic()

n2_histo
```

```{r}
o2_histo <- ggplot() +
  geom_histogram(data = o2_long,
                 aes(x = CONC, fill = TREATMENT),
                 binwidth = 2) +
  theme_classic()

o2_histo
```

```{r}
master_summary <- master %>%
  group_by(TREATMENT, TURBIDITY, TIME, ANALYTE) %>%
  summarise(mean = mean(CONC),
            se = sd(CONC)/sqrt(n()))
```

```{r}
n2_summary <- master_summary %>%
  filter(ANALYTE == "N2")

variance <- ggplot() +
  geom_point(data = n2_summary,
             aes(x = mean, y = se^2))

variance

o2_summary <- master_summary %>%
  filter(ANALYTE == "O2")

variance2 <- ggplot() +
  geom_point(data = o2_summary,
             aes(x = mean, y = se^2))

variance2




n2_raw_data <- ggplot() +
  geom_point(data = master %>%
               filter(ANALYTE == "N2"),
             aes(x = TURBIDITY, y = CONC, colour = TREATMENT),
             position = position_dodge(width = 5)) +
  facet_wrap(~TIME)

n2_raw_data

o2_raw_data <- ggplot() +
  geom_point(data = master %>%
               filter(ANALYTE == "O2"),
             aes(x = TURBIDITY, y = CONC, colour = TREATMENT),
             position = position_dodge(width = 5)) +
  facet_wrap(~TIME)

o2_raw_data
```


```{r}
plot2 <- ggplot() +
  geom_point(data = master_long %>%
               filter(analyte == "O2"),
             aes(x = turbidity, y = concentration, colour = sediment_source, size = day))

<<<<<<< HEAD
# ANOVA

```{r}
o2_anova_1 <- aov(data = o2_long %>%
                    filter(TREATMENT != "C"),
                  CONC ~ as.factor(TREATMENT) + as.factor(TURBIDITY) + as.factor(TIME))
=======
plot2
```


# ANOVA

```{r}
n2_anova_1 <- aov(data = n2_long,
                  concentration ~ as.factor(sediment_source)*as.factor(turbidity) + as.factor(day))
>>>>>>> 893a688b5ebf0b2a4d98d6032defc4c51acd4592

summary(o2_anova_1)
plot(o2_anova_1)
TukeyHSD(o2_anova_1)
```

```{r}
o2_anova_1 <- aov(data = o2_long %>%
                    filter(TREATMENT != "C"),
                  CONC ~ as.factor(TREATMENT)*as.factor(TURBIDITY)*as.factor(TIME))

summary(o2_anova_1)
plot(o2_anova_1)
TukeyHSD(o2_anova_1)
```

# GLM

Only additive effects.

```{r}
<<<<<<< HEAD
o2_glm <- glm(data = o2_long %>%
                    filter(TREATMENT != "C"),
              CONC ~ as.factor(TREATMENT) + TURBIDITY + TIME,
              family = gaussian(link = "identity"))
=======
o2_glm <- glm(data = o2_long,
              concentration ~ as.factor(sediment_source) + turbidity*day,
              family = gaussian(link = "log"))
>>>>>>> 893a688b5ebf0b2a4d98d6032defc4c51acd4592

summary(o2_glm)
plot(o2_glm)
```

<<<<<<< HEAD
=======







# Mixed-effect model

>>>>>>> 893a688b5ebf0b2a4d98d6032defc4c51acd4592
```{r}
o2_glm_2 <- glm(data = o2_long %>%
                    filter(TREATMENT != "C"),
              CONC ~ as.factor(TREATMENT)*TURBIDITY*TIME,
              family = gaussian(link = "identity"))

summary(o2_glm_2)
plot(o2_glm_2)
```


